---
title: "Run Datapack Model"
output: html_document
resource_files:
- "rsconnect/documents/datapack_model_job.Rmd/rstudio-connect.testing.ap.datim.org/flopez/datapack_model_job.dcf"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


The following chunk runs the datapack model on the posit server, this model is run and built with the `data-raw/model_calculations.R` file. This is just a wrapper, all changes are made to the `model_calculations.R` file. This RMD can be knit locally or run published on posit server. Note that when this report runs, it produces the capacity to download the datapack model run below.

```{r run_model, include=TRUE, echo=FALSE, message=FALSE, eval = T}

start_time <- as.POSIXlt(Sys.time())
paste0("Datapack model started ", start_time, " America/NY time")
source("model_calculations.R")
end_time <- as.POSIXlt(Sys.time())
paste0("Datapack model ended ", end_time, " America/NY time")
print(
  paste0(
    "Job run time: ",
    round(difftime(end_time, start_time, units = "mins"), 2),
    " mins"
  )
)

```


## Deltas, NA values, Which OUS and which indicators

if there are deltas below we see which ous and indicators there are deltas for:

``` {r ou_indicators, include = TRUE, echo = TRUE, message = TRUE, eval = T}

print(unique(deltas$ou))
print(unique(deltas$indicator_code))

# what types of indicators changed?
print(table(deltas_summary$indicator_type))


```

New NA values are summarized below (these are usually a sign something is wrong):

```{r nas, echo = T, include=T, message=T, eval = T}

# what about new na values?
new_nas <-
  deltas_summary %>%
  filter(count_new_nas > 0) %>%
  pull(indicator_code) %>%
  unique()

print(new_nas)

```

## Deltas Summary

if there are deltas below is a summary grouped by ou and indicator code with the indicator codes labeled as targets or results - DOWNLOAD FOR ANALYSIS WITH THE CSV BUTTON:

```{r analysis, include=TRUE, echo=FALSE, message=FALSE, eval = TRUE}

DT::datatable(
  deltas_summary,
  rownames = FALSE,
  filter = "top",
  extensions = "Buttons",
  options = list(
    dom = "Bftrip",
    buttons = c("csv")
  )
)

```


## Missing Indicator Combos

Below we mark which combos are IN the datapack schema BUT NOT IN the datapack model (missing combos are not necessarily an issue if they can be explained) - DOWNLOAD FOR ANALYSIS WITH THE CSV BUTTON:

``` {r missing_combos, include = TRUE, echo = FALSE, message = FALSE, eval = T}

DT::datatable(
  missing_schema_combos,
  rownames = FALSE,
  filter = "top",
  extensions = "Buttons",
  options = list(
    dom = "Bftrip",
    buttons = c("csv")
  )
)

```


Below we display the unique missing combo indicators from the above dataframe:

```{r echo = TRUE, include=TRUE, message=TRUE, eval = TRUE}

unique(missing_schema_combos$indicator_code)
```


## Download latest model

If there exist deltas between this model run and the latest production model then you can download the latest run below to validate and update in test s3 and prod:

```{r echo = FALSE, include=TRUE, message=TRUE, eval = TRUE}

if (NROW(deltas) > 0) {
  print("current server directory files: ")
  print(list.files())
  datapackcommons::flattenDataPackModel_21(cop_data) %>%
  downloadthis::download_this(
    output_name = new_dpm_file_name,
    output_extension = ".rds",
    button_label = "Download latest model run",
    button_type = "default",
    has_icon = TRUE,
    icon = "fa fa-save",
    class = "button_large"
  )
} else {
  print("No Differences in Model so no Model to Download")
}
```
