---
title: "Run PSNUXIM Model"
output: html_document
resource_files:
- "rsconnect/documents/psnuxim_model_job.Rmd/rstudio-connect.testing.ap.datim.org/flopez/psnuxim_model_job.dcf"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Run Script

The following chunk runs the PSNUXM model on the posit server, this model is run and built with the `data-raw/SnuxImDistMain.R` file. This is just a wrapper, all changes are made to the `SnuxImDistMain.R` file. This RMD can be knit locally or run published on posit server.

### TODO: new psnuxim model should be dumped into a staging folder in test S3, from there after it is deemed valid, it is moved to sharepoint and testing/prod S3 support_files directory for use in the apps.

```{r run_model, include=TRUE, echo=FALSE, message=FALSE}

start_time <- as.POSIXlt(Sys.time())
paste0("PSNUXIM model started ", start_time, " America/NY time")
source("SnuxImDistMain.R")
end_time <- as.POSIXlt(Sys.time())
paste0("PSNUXIM model ended ", end_time, " America/NY time")
print(
  paste0(
    "Job run time: ",
    round(difftime(end_time, start_time, units = "mins"), 2),
    " mins"
  )
)

```

## Deltas, NA values, Which OUS and which indicators

If there are deltas below we see which ous and indicators there are deltas for:

```{r echo = TRUE, include=TRUE, message=FALSE}

print(unique(deltas$ou))
print(unique(deltas$indicator_code))

```

New NA values are summarized below (these are usually a sign something is wrong):

``` {r analysis, echo = TRUE, include = TRUE, message = FALSE}

# what indicators values are newly na?
deltas_na <- deltas %>%
  filter(is.na(value.new))

# which rows and which unique indicators
print(nrow(deltas))
unique(deltas_na$indicator_code)

DT::datatable(
  deltas_na,
  rownames = FALSE,
  filter = "top",
  extensions = "Buttons",
  options = list(
    dom = "Bftrip",
    buttons = c("csv")
  )
)
```

## Missing data

Below we check if all the indicators in our MAPPING file are present in the new `PSNUXIM model`, if they aren't we should be able to explain why:

```{r echo=TRUE, include=TRUE, message=FALSE}

# are all indicators present?
setdiff(
   datapackcommons::Map24Tto25T$indicator_code,
   bind_rows(data) %>%
     pull(indicator_code) %>%
     unique()
 )
```
