---
title: "Run PSNUXIM Model"
output: html_document
resource_files:
- "rsconnect/documents/psnuxim_model_job.Rmd/rstudio-connect.testing.ap.datim.org/flopez/psnuxim_model_job.dcf"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


The following chunk runs the PSNUXM model on the posit server, this model is run and built with the `data-raw/SnuxImDistMain.R` file. This is just a wrapper, all changes are made to the `SnuxImDistMain.R` file. This RMD can be knit locally or run published on posit server. Note that when this report runs, it produces the capacity to download the psnuxim model run below.

```{r run_model, include=TRUE, echo=FALSE, message=FALSE}

start_time <- as.POSIXlt(Sys.time())
paste0("PSNUXIM model started ", start_time, " America/NY time")
source("SnuxImDistMain.R")
end_time <- as.POSIXlt(Sys.time())
paste0("PSNUXIM model ended ", end_time, " America/NY time")
print(
  paste0(
    "Job run time: ",
    round(difftime(end_time, start_time, units = "mins"), 2),
    " mins"
  )
)

```

## Deltas, NA values, Which OUS and which indicators

If there are deltas below we see which ous and indicators there are deltas for:

```{r echo = TRUE, include=TRUE, message=FALSE}

print(unique(deltas$ou))
print(unique(deltas$indicator_code))

```

New NA values are summarized below (these are usually a sign something is wrong):

``` {r analysis, echo = TRUE, include = TRUE, message = FALSE}

# what indicators values are newly na?
deltas_na <- deltas %>%
  filter(is.na(value.new))

# which rows and which unique indicators
print(nrow(deltas))
unique(deltas_na$indicator_code)

DT::datatable(
  deltas_na,
  rownames = FALSE,
  filter = "top",
  extensions = "Buttons",
  options = list(
    dom = "Bftrip",
    buttons = c("csv")
  )
)
```

## Missing data

Below we check if all the indicators in our MAPPING file are present in the new `PSNUXIM model`, if they aren't we should be able to explain why:

```{r echo=TRUE, include=TRUE, message=FALSE}

# are all indicators present?
setdiff(
   datapackcommons::Map24Tto25T$indicator_code,
   bind_rows(data_new) %>%
     pull(indicator_code) %>%
     unique()
 )
```


## Download latest model

If there exist deltas between this model run and the latest production model then you can download the latest run below to validate and update in test s3 and prod:

```{r echo = FALSE, include = TRUE, message = TRUE, eval = TRUE}

if (NROW(deltas) > 0) {
  print("current server directory files: ")
  print(list.files())
  data_new %>%
  downloadthis::download_this(
    output_name = new_psx_file_name,
    output_extension = ".rds",
    button_label = "Download latest model run",
    button_type = "default",
    has_icon = TRUE,
    icon = "fa fa-save",
    class = "button_large"
  )
} else {
  print("No Differences in Model so no Model to Download")
}
```
