---
title: "DataPackCommons Technical Notes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DataPackCommons Technical Notes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(datapackcommons)
library(tidyverse)
```


## What are the `DataPack model` and `PSNUxIM model` or `distribution` 

The `DataPack model` output is an `.rds` with historic target and result data used for building and validating DataPacks. See usage here [datapackr::packDataPack](https://github.com/pepfar-datim/datapackr/blob/11248b7138a98a4e1ba72e780ebf16ebb906ed61/R/packDataPack.R#L12) and here [datapackr::checkAnalytics](https://github.com/pepfar-datim/datapackr/blob/243cd2537661f5d1e9c340a9cffe1910ff8a483d/R/checkAnalytics.R#L562). Production versions of these models are stored in sharepoint and S3. The `DataPack model` includes historic data aggregating across mechanisms (including dedupes) to the PSNU level. So in the case of results this means we are aggregating from site and community data.

The `PSNUxIM model` output is an `.rds` file with historic target data broken out by mechanism. It is used for populating the PSNUxIM tab of the data pack with initial allocations of PSNU level targets to mechanisms including deduplication.

Recent versions of this files can be found on [sharepoint](https://pepfar.sharepoint.com/:f:/r/sites/PRIMEInformationSystems/Shared%20Documents/COP22%20Systems%20Support/Data%20Pack/Production%20Data%20Pack%20Models?csf=1&web=1&e=3EalWl). 

* A COP22 datapack model:
[model_data_pack_input_22_20220510_1_flat.rds](https://pepfar.sharepoint.com/:u:/r/sites/PRIMEInformationSystems/Shared%20Documents/COP22%20Systems%20Support/Data%20Pack/Production%20Data%20Pack%20Models/model_data_pack_input_22_20220510_1_flat.rds?csf=1&web=1&e=gnbfHy)

* A COP21 PSNUxIM distribution:
[PSNUxIM_COP21_2022-05-16](https://pepfar.sharepoint.com/:u:/r/sites/PRIMEInformationSystems/Shared%20Documents/COP22%20Systems%20Support/Data%20Pack/Production%20Data%20Pack%20Models/PSNUxIM_COP21_2022-05-16.rds?csf=1&web=1&e=GwuGdv)

* A COP22 PSNUxIM distribution:
[PSNUxIM_COP22_2022-05-16](https://pepfar.sharepoint.com/:u:/r/sites/PRIMEInformationSystems/Shared%20Documents/COP22%20Systems%20Support/Data%20Pack/Production%20Data%20Pack%20Models/PSNUxIM_COP21_2022-05-16.rds?csf=1&web=1&e=GwuGdv)

And here is a listing of the production support files on S3 as of 2022-05-18. Notice that we only keep the most current production version on S3, and the name remains stable. There is in the model generation scripts to post to S3 testing and prod environments. 

`aws-vault exec datapack-prod -- aws s3 ls prod.pepfar.data.datapack/support_files/`

`2021-02-04 23:36:17          0 `

`2022-05-10 14:26:18    3533160 datapack_model_data.rds`

`2022-05-16 21:15:17    4902501 psnuxim_model_data_21.rds`

`2022-05-16 18:58:52    5991842 psnuxim_model_data_22.rds`

## Generating a `DataPack model`

There is a script in `datapackcommons` `data-raw/model_calculations.R`. Commented out at the end of the script is valid code for writing the output locally and to S3. Currently the main output of the script is a nested data pack model, but this isn't what datapackr uses, so we must flatten this model using `flattenDataPackModel_21`. In the TODO we have an item to change this so that the flatten version is the base model and we stop creating a nested version. 

Note that in addition to the functions set up in the script, there are also functions included as part of the package. Ultimately all of the required functions should be moved over to datapackr as proper package functions.

There is also a function `diffDataPackModels` that allows you to do a diff between two different *flattened* DataPack models. This is very useful for refactoring because we can compare a full before and after version of the model to make sure code changes don't lead to (unexpected) model output changes.

Running the script start to finish (without uncommenting the writing of files), you will have a new version of the data pack model in an object called `cop_data`. This is the unflattened version which has some information useful in debugging, but this could be create separate from the data file itself.

## Generating a `PSNUxIM distribution/model`

There is a script in `datapackcommons` `data-raw/SnuxImDistMain.R`. It is currently possible to produce a PSNUxIM model for COP21 or COP22, and that variable `cop_year` is set at the top of the script. This script also contains code for comparing two versions of the PSNUxIM distribution, but it is inline at the bottom instead of in a function. It also contains code for sending the file to S3 commented out.

The the final R object created by this script is called data. When written as an RDS, these are the data used by datapackr to populate the PSNUxIM tab.

## Generating configuration files

There are 2 excel files used to configure the DataPack and PSNUxIM models.

1. `data-raw/model_calculations/model_calculations.xlsx`

    This contains a sheet with configuration details for pulling the historic data required for the DataPack (data_required sheet) as well as details on disaggregations required for the DataPack (dimension_item_sets sheet), which may include aggregation or disaggregation compared to what is actually in DATIM. I will provide more details on the configuration below.
1. `data-raw/snu_x_im_distribution_configuration/22Tto23TMap.xlsx`

    This workbook contains 1 sheet with the configuration details for generating the PSNUxIM distribution. more details on the configuration are provided below.
    
The three worksheets in these two excel files are saved in csv format for version control purposes, and these three csv files are converted to .rds files for inclusion on the package with the script `data-raw/CreateDataFiles.R`. This script includes some checks to ensure the internal consistency of the configuration (e.g. names and uids match) and it also provides some comparison of the data rds files to generated with the versions currently in the package. This allows us to ensure any configuration changes are as intended.

Practically what this means is that the process for updaing the DataPack model and the PSNUsxIM distribution model requires three things.

1. Update the relevant sheet(s) in `data-raw/model_calculations/model_calculations.xlsx` or `data-raw/snu_x_im_distribution_configuration/22Tto23TMap.xlsx` and save the xlsx file.

    NOTE: for next year the distribution file would be named  `23Tto24TMap.xlsx` if we keep the pattern the same.  
1. Save the updated Excel sheets as individual csv files
    - `data-raw/model_calculations/dimension_item_sets.csv`
    - `data-raw/model_calculations/data_required.csv`
    - `data-raw/snu_x_im_distribution_configuration/22Tto23TMap.csv`
1. Run the script that validates the confiuration files (csv versions) and creates the RDS files - `data-raw/CreateDataFiles.R`

## Configuring `dimension_item_sets`

You should familiarize yourself with the [DHIS2 API analytics endpoint](https://docs.dhis2.org/en/develop/using-the-api/dhis-core-version-238/analytics.html) in order to fully understand the content below.

The data in the DataPack is generally organized with an age and sex, or a key population in the rows (EID and GEND don't have row dissagregates, age is implied in EID columns). Additionally the columns often also have an implied disaggregation of some kind. For example `HTS_TST.KP.Pos.T_1` is disaggregated by key population status and an implied disagregation (analytics filer) on HIV test +. `TB_STAT.N.New.Pos.T_1` is disaggregated age: 1-65+, sex Male/Female and has an implied disagg on HIV Test Status (Specific) = Newly Tested Positives (Specific). This is similar for the PSNUxIM distribution.

The dimension item sets configuration is used to specify which disaggregations are used to pull data from the DHIS2 analytics endpoint for the DataPack Model and the PSNUxIM distribution. Each Column in the DataPack and unique indicator code in the PSNUxIM distributions can be associated with up to 3 dimension item or model sets. Age and/or sex and/or other disagg; or KP and/or other disagg.

An important aspect of the dimension item set configuration is it allows us to specify special situations for splitting historic data into new disaggregations for the datapack or aggregating historic data for use in the datapack. 

#### Examples

###### Standard key population disaggregation

This is the most basic example, we are pulling data from the `Key Populations v3` dimension, we include all of the dimension items under this dimension. As this dimension is actually a category, the category option uids and the dimension item uids match. The weights are all equal to 1 because we aren't aggregating or disaggeregating the data further for the data pack. The `model_set` her is `kp1` so that is the foreign key we will use in the other configuration files to pull data with these disaggregations. 

https://www.datim.org/api/dimensions?filter=name:eq:Key%20Populations%20v3&fields=:all,items[name,id]
https://www.datim.org/api/categories?filter=name:eq:Key%20Populations%20v3&fields=name,id,categoryOptions[name,id]

```{r}
dplyr::filter(datapackcommons::dim_item_sets, model_sets == "kp1") %>% as.data.frame()
```


###### Ages with additional distribution to new ages

Historically the highest age category has been 50+, but with COP22 and 23 the finer age bands are being use creating 50-54, 55-59, 60-64, and 65+. For the purposes of the datapack we need to be allocate or distribute historic data for the 50+ age category to these new age bands. Here is an example of a model set that does this:

```{r}
dplyr::filter(datapackcommons::dim_item_sets, model_sets == "15-65+") %>% as.data.frame()
```

`https://www.datim.org/api/dimensions?filter=name:eq:Age:%20Cascade%20Age%20bands&fields=:all,items[name,id]`


Some important things to note, this dimension does NOT come from a category, it is a a category group set. The dimension items are not necessarily category options, and the name and uids differ from category options. For instance the dimension item is `<item name="10-14 (Specific)" id="tIZRQs0FK5P"/>`, but the underlying category option is `<categoryOption name="10-14" id="jcGQdcpPSJP"/>`

https://www.datim.org/api/categoryOptionGroups/tIZRQs0FK5P?fields=:all,categoryOptions[name,id]

Also the 15-65+ model set does not include all of the dimension items from the `Age: Cascade Age bands` dimension/category group set.

In the configuration we can see that the 50+ dimension item appears 4 times in order to allocate the 50+ data to the 4 new age categories. This distribution of data happens in the data pack commons code. The 50+ data is proportionally distributed to 50-54, 55-59, 60-64, 65+ with weights .42, .35, .14, .09 (sum to 1), respectivly. 


Each `model_set` is a combination of one or more analytics dimensions and dimension items used in a DHIS2 analytics call. Most columns of dimension_item_sets are described here in the docs (`?datapackcommons::dim_item_sets`). The column `model_sets` in the dimension_items_sets excel sheet is semi-colon delimited and gets unnested in the package object `datapack::dim_item_sets`. A model set is the foreign key for the groupings that are used in the data_required and PSNUxIM model configurations to link the configuration to the ages, sexes, kps and other disaggs.

For every column of the datapack there will be one or more model sets

## Configuring a `PSNUxIM distribution`

## Configuring a `DataPack Model`

The DataPack model is used to populate certain columns of the `DataPack` with historic data in DATIM. Usually the target data for the current fiscal year/preceding COP year, and the results data from the prior fiscal year. So for COP 23 this means the historic data for the most part will be FY23/COP22 targets and FY22 results. The only edge case at the moment is `TB_PREV.N.R` which aggregates the last 5 years of data for COP22 (perhaps the prior 6 years for COP23).

In the datapack columns that take historic data are categorized as `mer/past` or `datapack/calculation`, e.g.

```{r}
head(dplyr::filter(datapackr::cop22_data_pack_schema, 
              (dataset == "mer" & col_type == "past") |
              (dataset == "datapack" & col_type == "calculation")))
```

Usually an indicator code with T_1 indicates the prior cop year (so T_1 in COP23 = FY23/COP22 targets = 2022Oct period) and R indicates results from the last fiscal year (so R in COP23 indicates FY22 results = 2021Oct period).

Consider the PMTCT tab of the datapack

Replicate vs distribute We never distribute 


## Key Files/Directories
```{r, echo = FALSE}
catalog <- tibble::tribble(~name, ~short_description,
                file.path("data-raw", "model_calculations.R"),
                "Script for generating datapack model",
                file.path("data-raw", "SnuxImDistMain.R"),
                "Script for generating PSNUxIM models",)
knitr::kable(catalog)
```


### TODO

* Modify GetData_DataPack to use datimutils::getAnalytics instead of internal getData_Analytics
* Modify SnuxImDistMain.R to use datimutils::getAnalytics instead of internal getData_Analytics,
* deprecate GetData_Analytics
* Consider removing period column in final output
* create flat file from the beginning and deprecate flatten COP function
* remove include military = true code, all countries now have military below country level as standard
* Make a function for comparing PSNUxIM distributions
* Create standard functions for sending stuff to S3
* Remove dimension item set model sets that are not used by the DataPackModel or the PSNUxIM distribution
* Begin analyzing indicators in the DataPack model that will be affected by new age categories
* Business as usual DataPack Model update
  - if datapack content remained the same, do we have the right indicators and disaggregations for pulling FY23 target data and FY22 results data as opposed to FY22 targets and FY21 results?
* Business as usual PSNUxIM distribution update
  - if PSNUxIM content remained the same, do we have the right standard dissagregations for pulling FY23 target data and maping to FY24 target data?
* Harmonize dimension_item_set model set names with the valida_age, valid_sex, and valid_kp names in the datapack template 
* Remove dim_cop_type row from dim_item_sets - we don't use it for anything
* enable tech_area, num_den, disagg_type in data_required asfilters. We use these in the SNUxIM distribution but not currently in `DataPack model`
  - `TB_PREV.N.R` condfiguration for this is quite awkward
* fully support mutli-item dimensions (with semicolons or pipes or some other special character)
  - added for `OVC_SERV.Active.T` and `OVC_SERV.Grad.T` PSNUxIM distribution but I think only partially implemented
* change full formula column into more of a description column perhaps with reference to data pack documentation

### Thoughts and ideas

* Reevealuate use desire to use indicators and basis of data pulls for `DataPack Model` - we don't use indicators for `PSNUxIM`
* 